---
title: "B.01 - Parsing ANNOVAR Results"
author: "Joshua H Cook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{B.01 - Parsing ANNOVAR Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Set up

Load libraries:

```{r load_libs, warning=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
library(tibble)
library(stringr)
library(KrasAlleleCna)
# devtools::load_all()
```

Establish paths:

```{r make_paths}
pkg_dir <- system.file(package = "KrasAlleleCna")
extdata_dir <- system.file("extdata", package = "KrasAlleleCna")
data_dir <- system.file("data", package = "KrasAlleleCna")
```

## Tidy data

'ANNOVAR' returned a separate file for each tissue sample (stored in "data-raw/gdc/annovar\_output"). In "data-raw/handle\_annovar.R", each file was read in as a tibble, and a list of tibbles was saved to "inst/extdata/".

```{r list_annovar}
# list of all annovar files
annovar_data <- readRDS(file.path(extdata_dir, "annovar_output_list.rds"))
```

Each file was read in as a tibble and the column `aa_mod` was added as a simpler version of the amino acid mutation. Only the rows with *KRAS* non-synonymous mutation data was kept.

```{r parse_annovar}
annovar_tib <- lapply(annovar_data, parse_annovar_mutation) %>%
    bind_rows(.id = "file_id") %>%
    mutate(file_id = str_remove_all(file_id, "\\.hg38_multianno\\.txt")) %>%
    filter(str_detect(Func.refGene, "exonic") &
           Gene.refGene == "KRAS" &
           !(ExonicFunc.refGene %in% c("synonymous SNV")))
head(annovar_tib)
```

A quick test to see if any mutations were missing.

```{r check_aamod}
any(is.na(annovar_tib$aa_mod))
```


## Save *tidy* data

```{r save_annovar}
usethis::use_data(annovar_tib, overwrite = TRUE)
```


### Show results

```{r show_annovar}
head(annovar_tib)
```

```{r kable_annovar}
kable(annovar_tib) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    scroll_box(width = "500px", height = "200px")
```
